\documentclass{article}[12pt]
\usepackage{fullpage}
\usepackage{epsfig}
\usepackage{amssymb}
\usepackage{amsmath}

\def\tf{{\hfill True \ \ \ False}}

\def\undshort{{\underline{\hskip1in}}}

\def\und2{{\underline{\hskip2in}}}

\begin{document}

{\bf\Large Final Exam} \hfill CSE 254 / Winter 2011
\\

\vspace{.25in}

Name: \underline{\hspace{3in}}
\\

ID: \underline{\hspace{3.2in}}
\\

\vspace{1in}

Be clear and concise. Write your answers in the space provided. Use the backs of pages for scratchwork.

\vspace{1in}

\begin{center}
\begin{tabular}{|c|c|} \hline
\hspace{1in} & \hspace{1in} \\
1 & \\
 &  \\ \hline
& \\
2 & \\
& \\ \hline
& \\
3 & \\
& \\ \hline
& \\
4 & \\
& \\ \hline
& \\
5 & \\
& \\ \hline
\end{tabular}
\end{center}

\pagebreak

\begin{enumerate}

\item Let $F$ be the set of functions defining disks in $R^2$,
  i.e. functions of the form 
\[
f(\vec{x}) = \begin{cases}
1 & \mbox{if } \|\vec{x}-\vec{c}\|_2 \leq r \cr
0 & \mbox{otherwise}
\end{cases}
\]
where $\vec{c} \in R^2$ is the center of the disk and $r\in R$ is it's
radius.

Find the VC dimension of $F$, prove your answer.

\pagebreak

\item
Let $F$ be the set of indicator function for second degree polynomials
on defines over $(x,y) \in R^2$, i.e. functions of the form
\[
f(x,y) = \begin{cases}
1 & \mbox{if } \sum_{0 \leq i+j \leq 2} a_{i,j} x^i y^j > 0 \cr
0 & \mbox{ otherwise}
\end{cases}
\]
give the tightest (smallest) upper bound you can on the number of
labelings for a set of $m$ instances.

\pagebreak

\item 

  Consider the problem of two people crossing an intersection from two
  different directions. The main source of conflict is that if both
  people cross at the same time, there is likely to be an accident,
  which will hurt both people.

  We can describe this interaction using the following game matrix
  between a row player and a column player (each entry defines row
  loss/column loss).
~\\
~\\
\begin{center}
\begin{tabular} {| c | c | c |}
\hline
 & cross & stop \\
\hline
cross & 100/100 & 0/5 \\
\hline
stop & 5/0 & 5/5 \\
\hline
\end{tabular}
\end{center}
~\\
~\\
\begin{enumerate}
\item
Find all of the Nash equilibria for this game and the expected loss of
each player under each of the equilibria.
\item
We say that an equilibrium is ``fair'' if the expected losses of all
the players are equal. One of the Nash equilibria for this game is
fair. Explain in words what behaviour of the two people is represented
by this equilibrium.
\item Find a {\em correlated} equilibrium that is fair and that has a
  lower expected loss for both players than the fair Nash equilibrium.
\end{enumerate}

\pagebreak
\item
  {\bf Question 4.6 in Cesa-Bianchi and Lugosi}\\
  Suppose your have $N$ experts whose loss at each iteration is a
  number in the range $[0,1]$. Show that the regret of ``follow the
  leader'' algorithm is bounded by the number of times the ``leader''
  (i.e. the action with minimal cumulative loss) changes during the
  sequence of plays.

\newcommand{\X}{{\cal X}}      % input space
\newcommand{\Mnk}{{\cal M}_{n,k}}
\newcommand{\Gn}{{\cal G}_{n}}
\newcommand{\braces}[1]{\left\{{#1}\right\}}
\newcommand{\paren}[1]{\left({#1}\right)}
\newcommand{\brackets}[1]{\left[{#1}\right]}
\newcommand{\sign}{{\rm sign}}
\newcommand{\pr}[2]{{\rm Pr}_{#1}\brackets{#2}} % Probability of event #2 according to distribution #1

\pagebreak
\item {\bf Question 3.6 in Schapire and Freund}\\
Let $\X_n=\{0,1\}^n$, and let $\Gn$ be a class of boolean functions
$g:\X_n\rightarrow\{-1,+1\}$.
Let $\Mnk$ be the class of all boolean functions that can be written
as a simple majority vote of $k$ functions in $\Gn$:
\[
  \Mnk \doteq \braces{f: x\mapsto \sign\paren{\sum_{j=1}^k g_j(x)}
                      \mid
                      g_1,\ldots,g_k\in \Gn}.
\]
In this problem, we will see, roughly speaking, that if $f$ can be
written as a majority vote of polynomially many functions in $\Gn$,
then under any distribution, $f$ can be approximated by some function
in $\Gn$.
But if $f$ cannot be so written as a majority vote, then there exists
some ``hard'' distribution under which $f$ cannot be approximated by
{\em any\/} function in $\Gn$.

\begin{enumerate}
\item
Show that if $f\in\Mnk$ then for all distributions $D$ on $\X_n$,
there exists a function $g\in\Gn$ for which
\[  \pr{x\sim D}{f(x)\neq g(x)} \leq \frac{1}{2} - \frac{1}{2k}. \]
\item
Show that if $f\not\in\Mnk$ then there exists a distribution $D$ on
$\X_n$ such that
\[
  \pr{x\sim D}{f(x)\neq g(x)} 
             > \frac{1}{2} - \sqrt{\frac{n\ln 2}{2k}}
\]
for every $g\in\Gn$.
\end{enumerate}

\end{enumerate}


\end{document}
