% $Header: /data/cvsroot/Courses/OnlineLearning/talks/talk4/talk4.tex,v 1.4 2006/01/24 08:08:47 yfreund Exp $

%\documentclass{beamer}
\documentclass[handout]{beamer}
% This file is a solution template for:

% - Giving a talk on some subject.
% - The talk is between 15min and 45min long.
% - Style is ornate.

% Copyright 2004 by Till Tantau <tantau@users.sourceforge.net>.
%
% In principle, this file can be redistributed and/or modified under
% the terms of the GNU Public License, version 2.
%
% However, this file is supposed to be a template to be modified
% for your own needs. For this reason, if you use this file as a
% template and not specifically distribute it as part of a another
% package/program, I grant the extra permission to freely copy and
% modify this file as you see fit and even to delete this copyright
% notice. 


\mode<presentation>
{
  \usetheme{Montpellier}

  %\setbeamercovered{transparent}
  % or whatever (possibly just delete it)
}

\usepackage{xmpmulti} % package that defines \multiinclude

\usepackage[english]{babel}

\usepackage[latin1]{inputenc}

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title[Online Bayes Alg.]% (optional, use only with long paper titles)
{Universal source coding \\ and \\ the Online Bayes algorithm}

\author[Freund] % (optional, use only with lots of authors)
{Yoav Freund}
% - Give the names in the same order as the appear in the paper.
% - Use the \inst{?} command only if the authors have different
%   affiliation.

\institute[Universities of Somewhere and Elsewhere] % (optional, but mostly needed)

\subject{Machine Learning}
% This is only inserted into the PDF information catalog. Can be left
% out. 

% If you have a file called "university-logo-filename.xxx", where xxx
% is a graphic format that can be processed by latex or pdflatex,
% resp., then you can add a logo as follows:

% \pgfdeclareimage[height=0.5cm]{university-logo}{university-logo-filename}
% \logo{\pgfuseimage{university-logo}}



% Delete this, if you do not want the table of contents to pop up at
% the beginning of each subsection:
%% \AtBeginSubsection[]
%% {
%%   \begin{frame}<beamer>
%%     \frametitle{Outline}
%%     \tableofcontents[currentsection,currentsubsection]
%%   \end{frame}
%% }


% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command: 

\beamerdefaultoverlayspecification{<+->}

\newcommand{\R}[1]{{\color{red}{#1}}}
\newcommand{\vp}{{\mathbf p}}
\newcommand{\HedgeLoss}{L_{\mbox{\footnotesize Hedge}}}

\input{../macros}

\begin{document}

%\iffalse %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
  \frametitle{Outline}
  \tableofcontents[pausesections]
  % You might wish to add the option [pausesections]
\end{frame}

\section{Combining experts in the log loss framework}

\begin{frame}
\frametitle{The log-loss framework}
\begin{itemize}
\item Algorithm \R{$A$} predicts a sequence \R{$c^1,c^2,\ldots, c^T$}
over alphabet \R{$\Sigma = \{1,2,\ldots,k\}$}
\item The prediction for the \R{$c^t$}th is a distribution over \R{$\Sigma$}:\\
\R{$\vp_A^t = \langle p_A^t(1),p_A^t(2),\ldots,p_A^t(k) \rangle$} 
\item When $c^t$ is revealed, the loss we suffer is \R{$-\log p_A^t(c^t)$}
\item The {\color{blue}cumulative log loss}, which we wish to minimize, 
is \R{$\lossalg^T = -\sum_{t=1}^T \log p_A^t(c^t)$}
\item \R{$\lceil \lossalg^T \rceil$} is the code length if \R{$A$} is combined with arithmetic coding.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{The game}
\begin{itemize}
\item Prediction algorithm \R{$A$} has access to \R{$N$} experts.
\item The following is repeated for \R{$t=1,\ldots,T$}
\begin{itemize}
\item Experts generate predictive distributions: \R{$\vp_1^t,\ldots,\vp_N^t$}
\item Algorithm generates its own prediction \R{$\vp_A^t$}
\item \R{$c^t$} is revealed.
\end{itemize}
\item {\bf Goal:} minimize regret:
\R{\[
-\sum_{t=1}^T \log p_A^t(c^t) + \min_{i=1,\dots,N} \paren{-\sum_{t=1}^T \log p_i^t(c^t)} 
\]}
\end{itemize}
\end{frame}

\section{The online Bayes Algorithm}

\begin{frame}
\frametitle{The online Bayes Algorithm}
\begin{itemize}
\item {\color{blue} Total loss} of expert \R{$i$}
\R{$$L_i^t = - \sum_{s=1}^{t} \log p_i^s(c^s);\;\;\; L_i^0 = 0$$}
\item {\color{blue}Weight} of expert \R{$i$}
\R{$$\wt{t}{i} = \wt{1}{i} e^{-L_i^{t-1}} = \wt{1}{i} \prod_{s=1}^{t-1} p_i^s(c^s)$$}
\item
Freedom to choose initial weights.\\
 \R{$\wt{1}{t} \geq 0$}, \R{$\sum_{i=1}^n \wt{1}{i} = 1$}
\item {\color{blue}Prediction} of algorithm \R{$A$}
\R{\[
\vp_A^t = \frac{\sum_{i=1}^N \wt{t}{i} \vp_i^t}{\sum_{i=1}^N \wt{t}{i}}
\]}
\end{itemize}
\end{frame}

\subsection{Comparison to \ouralg}

\begin{frame}
\frametitle{The \ouralg Algorithm}
Consider action \R{$i$} at time \R{$t$}
\begin{itemize}
\item Total loss:
\R{$$L_i^t = \sum_{s=1}^{t-1} \ell_i^s$$}
\item Weight:
\R{$$\wt{t}{i} = \wt{1}{i} e^{-\eta L_i^t}$$}
Note freedom to choose initial weight (\R{$\wt{1}{i}$})
\R{$\sum_{i=1}^n \wt{1}{i} = 1$}.
\item
\R{$\eta>0$} is the learning rate parameter. Halving: \R{$\eta \to \infty$}
\item Probability:
\R{$$\dist{t}{i} = \frac{\wt{t}{i}}{\sum_{j=1}^N \wt{t}{i}},\;\;
\pause     \distvec{t} = \frac{\wtvec{t}}{\sum_{j=1}^N \wt{t}{i}}$$}
\end{itemize}
\end{frame}

\section{The performance bound}
%\fi %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Cumulative loss vs. Final total weight}

\onslide<1-> Total weight: \R{$W^t \doteq \sum_{i=1}^N \wt{t}{i}$}

\onslide<2-> \R{$$
\frac{W^{t+1}}{W^t}  =  \frac{\sum_{i=1}^N \wt{t}{i} e^{\log p_i^t(c^t)}}{\sum_{i=1}^N \wt{t}{i}} 
\onslide<3->          =   \frac{\sum_{i=1}^N \wt{t}{i} p_i^t(c^t)}{\sum_{i=1}^N \wt{t}{i}} 
\onslide<4->        =  p_A^t(c^t)
$$}
\onslide<5-> \R{$$ -\log \frac{W^{t+1}}{W^t} = -\log p_A^t(c^t) $$}
\R{\[
\onslide<8-> -\log W^{T+1} =
\onslide<6-> -\log \frac{W^{T+1}}{W^1} = -\sum_{t=1}^T \log p_A^t(c^t)
\onslide<7-> = L_A^T
\]}
\onslide<9-> \R{\bf EQUALITY} not bound!
\end{frame}

\begin{frame}
\frametitle{Simple Bound}
\begin{itemize}
\item Use uniform initial weights \R{$\wt{1}{i} = 1/N$}
\item Total Weight is at least the weight of the best expert.
\R{\begin{eqnarray*}
L_A^T & = & -\log W^{T+1} 
\pause = -\log \sum_{i=1}^N \wt{T+1}{i} \\
\pause & = & -\log \sum_{i=1}^N \frac{1}{N} e^{-L_i^T} 
\pause  =  \log N - \log \sum_{i=1}^N e^{-L_i^T}\\
\pause & \leq & \log N - \log \max_i e^{-L_i^T}  
\pause = \log N + \min_i L_i^T
\end{eqnarray*}}
\item Dividing by $T$ we get
\R{$ \frac{L_A^T}{T} = \min_i \frac{L_i^T}{T} + \frac{\log N}{T} $}
\end{itemize}
\end{frame}

\subsection{Comparison to \ouralg}
\begin{frame}
\frametitle{Upper bound on \R{$\sumwts{\iter+1}$} for \ouralg}
\begin{lemma}[upper bound] 
For any sequence of loss vectors \R{$\costvec{1},\ldots,\costvec{\iter}$}
we have
\R{\[
\ln\paren{\sumwts{\iter+1}} \leq -(1-e^{-\eta}) \lossouralg.
\]}
\end{lemma}
\end{frame}

\begin{frame}
\frametitle{Tuning \R{$\eta$} as a function of \R{$T$}}
\begin{itemize}
\item trivially \R{$\min_i \lossi{i} \leq T$}, yielding
\R{\[
\lossouralg \leq \min_i \lossi{i} + \sqrt{2 T \ln N} + \ln N
\]}
\item per iteration we get:
\R{\[
\frac{\lossouralg}{T} \leq \min_i \frac{\lossi{i}}{T} + \sqrt{\frac{2 \ln N}{T}} + \frac{\ln N}{T}
\]}
\end{itemize}
\end{frame}

\subsection{Advantage over two part codes}

\begin{frame}
\frametitle{Bound better than for two part codes}
\begin{itemize}
\item
Simple bound as good as bound for two part codes (MDL) 
but enables online compression
\item Suppose we have \R{$K$} copies of each expert.
\item Two part code has to point to one of the $KN$ experts
\R{$ L_A \leq \log NK + \min_i L_i^T = \log NK + \min_i L_i^T$}
\item If we use Bayes predictor + arithmetic coding we get:
\R{$$ L_A = -\log W^{T+1} \leq \log  K \max_i \frac{1}{NK} e^{-L_i^T} 
=\log N + \min_i L_i^T $$}
\item We don't pay a penalty for copies.
\item More generally, the regret is smaller if many of the experts perform well.
\end{itemize}
\end{frame}


\section{Comparison with Bayesian Statistics}

\begin{frame}
\frametitle{Comparison with standard Bayesian statistics}
\begin{itemize}
\item The weight update rule is the same.
\item Normalized weights $=$ {\bf posterior probability  distribution}.
\item Bayesian analysis interested in the {\bf final} posterior.
\item Bayesian analysis assumes the data is {\bf generated} 
by a distribution in the support of the prior.
\item Goal of Bayesian is to {\bf estimate true distribution}, 
goal of online learning is to {\bf minimize regret}.
\item Optimality of algorithm is {\bf axiom} of Bayesian statistics.
\item Bayesian methods perform poorly when the loss is not log loss
  and the data not generated by a distribution in the support.
\begin{itemize}
\item
Loss can sometimes be defined through the noise distribution: square loss is equivalent to assuming guassian noise.
\item
For number of mistakes - Bayesian method cannot be ``fixed''. Requires variable learning rate.
\end{itemize}
\end{itemize}
\end{frame}

% \section{Least Informative Priors}

\section{Computational issues}

\begin{frame}
\frametitle{Computational Issues}
\begin{itemize}
\item Naive implementation: calculate the prediction of each of 
the \R{$N$} experts.
\item Puts severe limit on number of experts.
\item What if set of experts is uncountably infinite.
\item Bayesian tricks:
\begin{itemize}
\item {\bf Conjugate priors}: A prior over a continuous domain whose functional form does not change with when updated. 
\pause Number of parameters defining posterior is constant. 
\pause Update rule translates into update of parameters.
\pause parameters correspond to ``sufficient statistics''.
\pause exists for the familty of exponential distributions.
\item {\bf Markov Chain Monte Carlo} Sample the posterior. 
\pause Can sometimes be done efficiently.
\pause Efficient sampling relates to mixing rate of markov chain whose limit dist is the posterior dist. 
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Next class}
\begin{itemize}
\item How to deal with an uncountably infinite class of models.
\item To maintain {\color{red} S}atisfactory status in class:
\begin{itemize}
\item Register on TWiki and update your information.
\item If you are taking the class for 4 points, start a project page.
\end{itemize}
\item For {\color{red} EXTRA} credit
\begin{itemize}
\item Post (and answer) questions.
\item Read the background material I put on the class twiki page (for
  class no. 5)
\end{itemize}
\end{itemize}
\end{frame}

\iffalse %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{XXX}
\begin{itemize}
\item XXX
\end{itemize}
\end{frame}

\fi %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}


